Spatial Analysis

### 2022-09-20 Scripts/spatial_analisis.Rmd
Updated script obe reporductible

```{r setup, include=FALSE}
library(MyFunctions)
MyFunctions::my_lib(c("ggmap","sf","st","tidyverse","tools","readr","data.table","maps","viridis","wesanderson","knitr","kableExtra","plotly","ggrepel","ggsflabel","janitor","interp","rnaturalearth","readxl","cowplot"))

@@

# Include figure for paper

land_sf$group <- factor(state_mpa$abrev,      # Reordering group factor levels
                          levels = state_lat$abrev)

ggplot() +
  geom_sf(data = st_as_sf(map("state", plot = FALSE, fill = TRUE))) +
  geom_sf(data = land_sf, aes(fill = state)) +
  scale_fill_manual(values = state_pallet) +
  my_ggtheme_m(leg_pos = "right")

@@

x <- as.data.frame(st_read(my_path("D", "Partial/grid_fp",name="grid_eez_fp.shp"))) %>% 
  select(index,spp = spcs_nm,port_name = port_nm, abrev = stt_pst) %>% 
  bind_cols(
    as.data.frame(st_coordinates(st_read(my_path("D", "Partial/grid_fp",name="grid_eez_fp.shp"))))
  ) %>% 
  rename(lon = X, lat = Y) %>% 
  left_join(as.data.frame(land_sf),
            by = "abrev") %>% 
  mutate(state= str_to_sentence(ID)) %>% 
  select(1:6,state) %>% 
  mutate(
    spp = ifelse(spp == "summer flounder","Paralichthys dentatus", 
                 ifelse(spp == "scup", "Stenotomus chrysops", 
                        "Centropristis striata"
                 )
    )
  )

write_csv(x,my_path("D", "Partial/grid_fp",name="grid_eez_fp_df.csv"))

@@

grids_sw <- bind_rows(
  my_path("G", "AcrossBoundaries/Data/Partial/grid_sw", name = "grid_eez_sw_df.csv", read = T) %>% 
    mutate(spp = "Paralichthys dentatus"),
  my_path("G", "AcrossBoundaries/Data/Partial/grid_sw", name = "grid_eez_sw_df.csv", read = T) %>% 
    mutate(spp = "Stenotomus chrysops"),
  my_path("G", "AcrossBoundaries/Data/Partial/grid_sw", name = "grid_eez_sw_df.csv", read = T) %>% 
    mutate(spp = "Centropristis striata")
) %>% 
  left_join(land_sf) %>% 
  select(index,state,abrev,lat,lon,spp) %>% 
  mutate(spatial = "sw",
         port_name = NA
  )

grids_fp <- my_path("G", "AcrossBoundaries/Data/Partial/grid_fp", name = "grid_eez_fp_df.csv", read = T
) %>%
  mutate(spatial = "fp",
         spp = ifelse(spp == "summer flounder","Paralichthys dentatus", 
                      ifelse(spp == "scup", "Stenotomus chrysops", 
                             "Centropristis striata"
                      )
         )
  )

@@

# Ports data
ports_df <- my_path("G","AcrossBoundaries/Data/Partial", "top_ports.csv", read = T) %>% 
  filter(tresh == "25th") %>%
  group_by(species_name,state_postal) %>% 
  summarise(n =n()) %>% 
  select(-n)
# distinct(c(,state_postal), .keep_all = T) # No need for duplication

@@

spp_paths <- list.files(my_path("G",extra_path = "AcrossBoundaries/Results/Partial/Interpolation/"))

# Some gibiris for names
taxon_list <- gsub("_"," ",spp_paths)
taxon_list <- gsub("tif ","",taxon_list)
taxon_list <- gsub(".csv","",taxon_list)

taxon_list <- taxon_list[taxon_list %in% Species]

# Now set a list of files to read
taxon_read <- paste0("tif_",taxon_list,".csv")
taxon_read <- gsub(" ","_",taxon_read)

taxon_read <- spp_paths[spp_paths %in% taxon_read]

@@

taxon_sw <- my_path("G",extra_path = "AcrossBoundaries/Results/Partial/Interpolation/", name = taxon_read)

# Load all sp in one table
historic_spp <- bind_rows(lapply(taxon_sw, fread)) %>% 
  left_join(grids,
            by = c("index","lon","lat","spp")
  ) %>%
  filter(!is.na(spatial)) %>% 
  mutate(
    season = ifelse(str_detect(region,"Spring"),"Spring","Fall"),
    label = ifelse(spp == "Paralichthys dentatus" & year >= 1980 & year <= 1986,"Reference",
                   ifelse(spp == "Stenotomus chrysops" & year >= 1988 & year <= 1992,"Reference",
                          ifelse( spp == "Centropristis striata" & year >= 1980 & year <= 2001,"Reference",
                                  ifelse(year > 2009,"2010-2019",NA)
                          )
                   )
    )
  )

@@

# re order states by latitude
land_sf_t <- land_sf

land_sf_t$group <- factor(land_sf_t$abrev,      # Reordering group factor levels
                               levels = state_lat$abrev)

 land_sf_t <- land_sf_t %>% 
  left_join(state_lat) %>% 
   mutate(
     name = paste0(str_to_title(state)," (",abrev,")")
   )

@@

 # Plot figure
study_plot <-
  ggplot() +
  geom_sf(data = us_map, aes(),fill = "grey", color = "lightgrey") +
  geom_sf(data = land_sf_t, aes(fill = reorder(name,-order))) +
  scale_fill_manual("State",
                    values = state_pallet) +
  my_ggtheme_m(leg_pos = "right",map_type = "R")

@@

my_path("D","Partial","top_ports.csv", read = T) %>% 
  filter(tresh == "25th",
         state_postal == "NY") %>% 
  ggplot() +
  geom_point(
    aes(
      x = lon,
      y = lat,
      color = port_name
    )
  )

@@

# Load grid of state waters
grid_eez_sw_sf <-  st_read(my_path("G","AcrossBoundaries/Data/Partial/grid_sw", name = "grid_eez_sw_sf.shp")) %>%
  select(index,state,geometry) %>% 
  mutate(method = "state_waters",
         state = str_to_sentence(state)) %>% 
  left_join(state_lat)

@@

# Load grid of fishing ports
grid_eez_fp_sf <-  st_read(my_path("G","AcrossBoundaries/Data/Partial/grid_fp", name = "grid_eez_fp.shp")) %>%
  select(index,abrev,geometry) %>% 
  mutate(method = "fishing_port"#,
         # state = str_to_sentence(landing_port)
  ) %>% 
  left_join(state_lat)

@@

top_ports_sen <- my_path("G","AcrossBoundaries/Data/Partial","top_ports.csv", read = T)

@@

sens_plot <- grid_fp_bf_dt %>% 
  group_by(species_name,state_postal,Threshold = tresh) %>% 
  summarise(n = n(),
            nuni = length(unique(as.character(index)))
  ) %>% 
  ggplot() +
  geom_bar(
    aes(
      x = Threshold,
      y = nuni,
      fill = Threshold
    ),
    stat = "identity"
  ) +
  facet_grid(species_name~state_postal) +
  my_ggtheme_p(leg_pos = "right",
               ax_tx_s = 11
               ) +
  theme(axis.text.x = element_blank(),
        strip.background = element_rect(fill = "grey"))+
  labs(y = "Number of grid-cells (n)", x = "Threshold") +
  scale_fill_brewer(palette = "Set3")

@@

```
# Old code

>>>>

```{r setup, include=FALSE}
library(MyFunctions)
MyFunctions::my_lib(c("ggmap","sf","st","tidyverse","tools","readr","data.table","maps","viridis","wesanderson","knitr","kableExtra","plotly","ggrepel","ggsflabel","janitor","interp","rnaturalearth","readxl","cowplot","here"))

@@

@@

x <- as.data.frame(st_read(here("Data/Processed/grid_fp/grid_eez_fp.shp"))) %>% 
  select(index, spp, port_name, abrev) %>% 
  bind_cols(
    as.data.frame(st_coordinates(st_read(here("Data/Processed/grid_fp/grid_eez_fp.shp"))))
  ) %>% 
  rename(lon = X, lat = Y) %>% 
  left_join(as.data.frame(land_sf),
            by = "abrev") %>% 
  mutate(state= str_to_sentence(ID)) %>% 
  select(1:6,state) %>% 
  mutate(
    spp = ifelse(spp == "summer flounder","Paralichthys dentatus", 
                 ifelse(spp == "scup", "Stenotomus chrysops", 
                        "Centropristis striata"
                 )
    )
  )

# write_csv(x,my_path("D", "Partial/grid_fp",name="grid_eez_fp_df.csv"))

@@

grids_sw <- bind_rows(
  read_csv(here("Data/Processed/grid_sw/grid_eez_sw_df.csv")) %>% 
    mutate(spp = "Paralichthys dentatus"),
  read_csv(here("Data/Processed/grid_sw/grid_eez_sw_df.csv")) %>% 
    mutate(spp = "Stenotomus chrysops"),
  read_csv(here("Data/Processed/grid_sw/grid_eez_sw_df.csv")) %>% 
    mutate(spp = "Centropristis striata")
) %>% 
  left_join(land_sf) %>% 
  select(index,state,abrev,lat,lon,spp) %>% 
  mutate(spatial = "sw",
         port_name = NA
  )

grids_fp <- read_csv(here("Data/Processed/grid_fp/grid_eez_fp_df.csv")) %>%
  mutate(spatial = "fp",
         spp = ifelse(spp == "summer flounder","Paralichthys dentatus", 
                      ifelse(spp == "scup", "Stenotomus chrysops", 
                             "Centropristis striata"
                      )
         )
  )

@@

# Ports data
ports_df <- read_csv(here("Data/Processed/top_ports.csv")) %>% 
  filter(tresh == "75th") %>% # this selects the top 25%
  group_by(species_name,state_postal) %>% 
  summarise(n =n()) %>% 
  select(-n)
# distinct(c(,state_postal), .keep_all = T) # No need for duplication

@@

# Get files
spp_paths <- list.files(here("Data/Processed/Interpolation"),
                        full.names = T)

@@

# Load all sp in one table
historic_spp <- bind_rows(lapply(spp_paths, fread)) %>% 
  left_join(grids,
            by = c("index","lon","lat","spp")
  ) %>%
  filter(!is.na(spatial)) %>% 
  mutate(
    season = ifelse(str_detect(region,"Spring"),"Spring","Fall"),
    label = ifelse(spp == "Paralichthys dentatus" & year >= 1980 & year <= 1986,"Reference",
                   ifelse(spp == "Stenotomus chrysops" & year >= 1988 & year <= 1992,"Reference",
                          ifelse( spp == "Centropristis striata" & year >= 1980 & year <= 2001,"Reference",
                                  ifelse(year > 2009,"2010-2019",NA)
                          )
                   )
    )
  )

@@

# re order states by latitude
land_sf$group <- factor(land_sf$abrev,      # Reordering group factor levels
                               levels = state_lat$abrev)

 land_sf_to_plot <- land_sf %>% 
  left_join(state_lat) %>% 
   mutate(
     name = paste0(str_to_title(state)," (",abrev,")")
   )

@@

 # Plot figure 0 ( NewFig 1)
study_plot <- ggplot() +
  geom_sf(data = us_map, aes(),fill = "grey", color = "lightgrey") +
  geom_sf(data = land_sf_to_plot, aes(fill = reorder(name,-order))) +
  scale_fill_manual("State",
                    values = state_pallet) +
  my_ggtheme_m(leg_pos = "right",map_type = "R")

@@

read_csv(here("Data/Processed/top_ports.csv")) %>% 
  filter(tresh == "75th",
         state_postal == "NY") %>% 
  ggplot() +
  geom_point(
    aes(
      x = lon,
      y = lat,
      color = port_name
    )
  )

@@

# Load grid of state waters
grid_eez_sw_sf <-  st_read(here("Data/Processed/grid_sw/grid_eez_sw_sf.shp")) %>%
  select(index,state,geometry) %>% 
  mutate(method = "state_waters",
         state = str_to_sentence(state)) %>% 
  left_join(state_lat)

@@

# Load grid of fishing ports
grid_eez_fp_sf <-  st_read(here("Data/Processed/grid_fp/grid_eez_fp_sf.shp")) %>%
  select(index,abrev,geometry) %>% 
  mutate(method = "fishing_port"#,
         # state = str_to_sentence(landing_port)
  ) %>% 
  left_join(state_lat)

@@

top_ports_sen <- read_csv(here("Data/Processed/top_ports.csv"))

@@

sens_plot <- grid_fp_bf_dt %>% 
  group_by(species_name,state_postal,Threshold = tresh) %>% 
  summarise(n = n(),
            nuni = length(unique(as.character(index)))
  ) %>% 
  ggplot() +
  geom_bar(
    aes(
      x = Threshold,
      y = nuni,
      fill = Threshold
    ),
    stat = "identity"
  ) +
  facet_grid(species_name~state_postal) +
  my_ggtheme_p(leg_pos = "right",
               ax_tx_s = 11
               ) +
  theme(axis.text.x = element_blank(),
        strip.background = element_rect(fill = "grey"))+
  labs(y = "Number of grid-cells (n)", x = "Top Ports") +
  scale_fill_brewer(palette = "Set3")

@@

# Old code
